---
title: "vignette_deconv"
author: "Clémentine Decamps"
date: "6 juin 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



#Data


```{r datas}
D = readRDS("/home/decampcl/projects/DNAdeconv/data/D.rds")
dim(D)
head(D)[1:5, 1:5]
exp_grp = readRDS("/home/decampcl/projects/DNAdeconv/data/exp_grp.rds")
dim(exp_grp)
head(exp_grp)[1:5, 1:5]
```


#Step 1 : filter 

We want to remove probes correlated with confounders variables. For this, we used the function `CF_detection` which makes a linear regression.

```{r CF}
D_CF = DNAdeconv::CF_detection(D, exp_grp)

dim(D_CF)

print(paste0("Number of correlated probes removed : ", nrow(D) - nrow(D_CF)))
```

#Step 2 : choose k 

To choose the number k of cell types, we used the function `plot_k` which makes a PCA.

```{r choice_k}
DNAdeconv::plot_k(D_CF)  
```
Here with the Cattel's rule we select 4 principal components, that is to say 5 cell types.

#Step 3 : feature selection 

We also want te select a subset of appropriate features. For this, we used `feature_selection` which keep only most variable probes.

```{r FS}
D_FS = DNAdeconv::feature_selection(D_CF, pval = 0.02)

dim(D_FS)

print(paste0("Number of low variance probes removed : ", nrow(D_CF) - nrow(D_FS)))
```

#Step 4 : deconvolution

There is 3 methods of deconvolution.

```{r deconv}
results_RFE = DNAdeconv::RFE(D_FS, nbcell = 5)
results_MDC = DNAdeconv::MDC(D_FS, nbcell = 5, lambdas = c(0,10^(-5:-1)))

infloci = read.table("/home/decampcl/projects/DNAdeconv/data/inf_loci.txt", header = FALSE, sep = "\t")
results_Edec = DNAdeconv::Edec(D_FS, nbcell = 5, infloci = infloci)
```
